---
title: 'STAT 231: Problem Set 2B'
author: "Jack Dove"
date: "due by 5 PM on Friday, September 11"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Series B homework assignments are designed to help you futher ingest and practice the material covered in class over the past week(s).  You are encouraged to work with other students, but all code must be written by you and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps2B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps2B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(mdsr)   
library(tidyverse)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 

# If you discussed this assignment with any of your peers, please list who here:

> ANSWER:

\newpage
# MDSR Exercise 4.14 (modified)

Use the `Pitching` data frame from the `Lahman` package to identify every pitcher in baseball history who has accumulated at least 300 wins (`W`) and at least 3,000 strikeouts (`SO`).  

a. How many pitchers meet this criteria?  

> ANSWER: 10 pitchers. 

```{r}
library(Lahman)
Pitching1<- Pitching %>%
  select(playerID, yearID, W, SO)

Pitching1 %>%
  group_by(playerID) %>%
  summarize(
    totalW = sum(W), totalSO = sum(SO)
  ) %>%
  filter(totalW >299 & totalSO > 2999) %>% 
  arrange(desc(totalSO))
```

b. Which of these pitchers had the most accumulated strikeouts?  How many strikeouts had he accumulated?  What is the most strikeouts he had in one season?

> ANSWER: Nolan Ryan accumulated 5714 total strikeouts. In 1973, he struck out (wow) 383 batters. 

```{r}
Pitching2 <- Pitching1 %>%
  filter(playerID =="ryanno01") %>%
  arrange(desc(SO))
Pitching2
```


\newpage 
# MDSR Exercise 4.17 (modified)

a. The Violations data set in the `mdsr` package contains information regarding the outcome of health inspections in New York City.  Use these data to calculate the median violation score by zipcode and dba for zipcodes in Manhattan.  What pattern (if any) do you see between the number of inspections and the median score?  Generate a visualization to support your response.  

> ANSWER: I see a medium positive relationship between number of inspections and median violation score by zipcode and by DBA. I generated a dual-facet smoothed scatterplot, which, using the log of inspections, shows an initial strong positive relationship between the two variables, then tapers off at about 9 total inspections. These results are also supported by moderate correlation coefficients, both between 0.4-0.6 and suggesting a medium link between log n inspections and median score for both categories. 

```{r}
data(Violations)
head(Violations, 10)

#By Zipcode
Violations1 <- Violations %>%
  filter(is.na(score)==FALSE) %>%
  group_by(zipcode) %>%
  summarise(medianscore = median(score), ninspections = log(n()))
Violations1
plot1<-ggplot(data=Violations1) +
  geom_point(mapping = aes(x = ninspections, y = medianscore)) + 
  geom_smooth(mapping= aes(x = ninspections, y = medianscore), type = "loess") +
  labs(
    x="Log Number of Inspections",
    y="Median Violation Score",
    title= "NYC DBAs"
)
  

#By DBA
Violations2 <- Violations %>%
  filter(is.na(score)==FALSE) %>%
  group_by(dba) %>%
  summarise(medianscore = median(score), ninspections = log(n())) 
Violations2
plot2 <- ggplot(data=Violations2) +
  geom_point(mapping = aes(x = ninspections, y = medianscore)) + 
  geom_smooth(mapping= aes(x = ninspections, y = medianscore), type = "loess") +
  labs(
    x="Log Number of Inspections",
    y="Median Violation Score",
    title= "NYC Zipcodes"
)

library(cowplot)
#Put the charts together
plot_grid(plot1, plot2) 

cor(medianscore~ninspections, data=Violations1)
cor(medianscore~ninspections, data=Violations2)
```

b.  In your visualization in part (a), there should be at least a few points that stand out as outliers.  For *one of the outliers*, add text to the outlier identifying what business it is and an arrow pointing from the text to the observation.  First, you may want to `filter` to identify the name of the business (so you know what text to add to the plot).  

(Can't remember how to create a curved arrow in `ggplot`?  Can't remember how to add text to the plot in `ggplot`?  Check out the answers to questions #5 and #8, respectively, in the Moodle R Q&A forum!)

```{r}
outlier <- Violations2 %>%
  filter(medianscore > 110)
outlier

plot2 + geom_curve(x = 7, xend = 3.3, y = 50, yend = 112, arrow = arrow(length = unit(0.3,"cm")), curvature = 0.5) + geom_text(x=7, y=50, label= "Roxy Diner")
```


\newpage
# MDSR Exercise 5.7

Generate the code to convert the data frame shown with this problem in the textbook (on page 130, and shown below) to wide format (i.e., the result table).  Hint: use `gather()` in conjuction with `spread()`; OR `pivot_longer()` in conjuction with `pivot_wider()`.


```{r}
FakeDataLong <- data.frame(grp = c("A","A","B", "B")
                           , sex = c("F", "M", "F", "M")
                           , meanL = c(0.22, 0.47, 0.33, 0.55)
                           , sdL = c(0.11, 0.33, 0.11, 0.31)
                           , meanR = c(0.34, 0.57, 0.40, 0.65)
                           , sdR = c(0.08, 0.33, 0.07, 0.27))
# Looked up unite function on internet 
Table <- FakeDataLong %>%
  gather(key = "x", value = "y", meanL, meanR, sdL, sdR) %>%
  unite(col = "x1", sex, x, sep = ".") %>%
  spread(x1, y)
Table
```

\newpage
# PUG Post

What topics or questions are you interested in exploring related to your PUG theme?  Dream big here.  Don't worry about whether there is data out there that's available and accessible that you could use to address your questions/topics.  Just brainstorm some ideas that get you excited.  In your PUG team discussion forum on GitHub, start a thread called "Brainstorming" (or, if another team member has already started the thread, reply to their post) with your ideas.

> ANSWER: Do not write anything here.  Write down your ideas in your PUG team's discussion thread titled "Brainstorming" on GitHub.
