---
title: "STAT 231: Problem Set 3B"
author: "Jack Dove"
date: "due by 5 PM on Friday, September 18"
output: pdf_document
---

This homework assignment is designed to help you futher ingest, practice, and expand upon the material covered in class over the past week(s).  You are encouraged to work with other students, but all code and text must be written by you, and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps3B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps3B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 
## If you discussed this assignment with any of your peers, please list who here:

> ANSWER:

\newpage
## Shiny app

1. Finish your app from Lab04b and add your app code to the R code chunk below:

(1) update the Lab04b app to still explore the `hate_crimes` dataset, but with different app functionality (e.g. different widgets, variables, layout, theme...); OR
(2) use it as a template to create a Shiny app for a different dataset, choosing from:

* `mad_men` (tv performers and their post-show career)
* `ncaa_w_bball_tourney` (women's NCAA div 1 basketball tournament, 1982-2018)
* `nfl_suspensions` (NFL suspensions, 1946-2014)
* `candy_rankings` (candy characteristics and popularity)

These four datasets are also part of the `fivethirtyeight` package and their variable definitions are included in a pdf posted to the Moodle course page.

If using the `hate_crimes` dataset, be sure to update:

* at least 2 different widgets; and
* the layout (e.g. not in tabs or different page layout) or the theme
     + check out:  https://rstudio.github.io/shinythemes/
* like a challenge?  incorporate one of the click, hover or brush features
     + check out: https://shiny.rstudio.com/articles/plot-interaction.html

      
```{r, eval=FALSE}
library(fivethirtyeight)
library(shinythemes)
library(tidyverse)
library(datasets)
library(janitor)
library(DT)
library(mosaic)
library(readxl)
library(ICON)
library(leaflet)
library(GISTools)
library(shiny)
library(dplyr)


#NFL TEAM Locations (note: the file path was removed in published version)
stadiums_nfl <- read_excel("/Users/jackdove/Desktop/Data Science/Stat231JackDove/Labs/Lab04b-shiny-nflsuspensions/stadiums_nfl.xlsx") 
stadiums_nfl1 <- stadiums_nfl %>%
  dplyr::select(ABBREVIATION, TEAM, LONGITUDE, LATITUDE, CONFERENCE, DIVISION) %>%
  mutate(team = ABBREVIATION) 


#Add locations to suspensions data set
nfl_suspensions1 <- nfl_suspensions %>%
  filter(team != "FREE") %>%
  filter(games != "Indef.") %>%
  group_by(team) %>%
  summarize(NumSuspensions = n()) %>%
  inner_join(stadiums_nfl1, by=c("team" = "team")) 


                               
division_choices <- (nfl_suspensions1 %>%
                         count(DIVISION))$DIVISION  
conference_choices <- (nfl_suspensions1 %>%
                         count(CONFERENCE))$CONFERENCE  

ui <- fluidPage(
  
  h1("Team-based Overview of NFL Player Suspensions"),
  
  sidebarLayout(
    sidebarPanel(
      checkboxGroupInput(inputId = "div"
                         , label = "Include divisions: "
                         , choices = division_choices
                         , selected = division_choices
                         , inline = TRUE),
      checkboxGroupInput(inputId = "conf"
                         , label = "Include conferences: "
                         , choices = conference_choices
                         , selected = conference_choices
                         , inline = TRUE),
    ),
    mainPanel (
    navbarPage(title = "NFL Suspensions",
               tabPanel(title = "Team Comparison",
                        plotOutput("scatter")),
               tabPanel("Map",
                        leafletOutput("map")
                        )         
    )
    )
  )
)

server <- function(input,output){
  
  use_data <- reactive({
    data <- filter(nfl_suspensions1, DIVISION %in% input$div & CONFERENCE %in% input$conf)
  })
  
  
  output$scatter <- renderPlot({
    ggplot(data = use_data(), aes(x = TEAM, y = NumSuspensions)) +
      geom_col() +
      labs(y = "Number of Suspensions", title = "NFL Suspensions 1946-2014 by Team")  +
      theme(axis.text.x = element_text( 
        angle = 90, 
        color="blue", 
        size=15, 
        face=3)
      )
  })
  
  output$map <- renderLeaflet({
    
    #TEAM Icons
    NFL_icons <- list(
      iconUrl =       ifelse(use_data()$TEAM=="Pittsburgh Steelers", "http://prod.static.steelers.clubs.nfl.com/assets/images/svg/SteelersMark.svg",
                             ifelse(use_data()$TEAM=="Cleveland Browns","https://static.nfl.com/static/site/img/logos/svg/teams/CLE.svg",
                                    ifelse(use_data()$TEAM=="Cincinnati Bengals","https://static.nfl.com/static/site/img/logos/svg/teams/CIN.svg",
                                           ifelse(use_data()$TEAM=="Baltimore Ravens","https://static.nfl.com/static/site/img/logos/svg/teams/BAL.svg",
                                                  ifelse(use_data()$TEAM=="Chicago Bears","https://static.nfl.com/static/site/img/logos/svg/teams/CHI.svg",
                                                         ifelse(use_data()$TEAM=="Houston Texans","https://static.nfl.com/static/site/img/logos/svg/teams/HOU.svg",
                                                                ifelse(use_data()$TEAM=="Kansas City Chiefs","https://static.nfl.com/static/site/img/logos/svg/teams/KC.svg",
                                                                       ifelse(use_data()$TEAM=="New Orleans Saints","https://static.nfl.com/static/site/img/logos/svg/teams/NO.svg",
                                                                              ifelse(use_data()$TEAM=="Miami Dolphins","https://static.nfl.com/static/site/img/logos/svg/teams/MIA.svg",
                                                                                     ifelse(use_data()$TEAM=="Tampa Bay Buccaneers","https://static.nfl.com/static/site/img/logos/svg/teams/TB.svg",
                                                                                            ifelse(use_data()$TEAM=="Atlanta Falcons","https://static.nfl.com/static/site/img/logos/svg/teams/ATL.svg",
                                                                                                   ifelse(use_data()$TEAM=="Buffalo Bills","https://static.nfl.com/static/site/img/logos/svg/teams/BUF.svg",
                                                                                                          ifelse(use_data()$TEAM=="San Francisco 49ers","https://static.nfl.com/static/site/img/logos/svg/teams/SF.svg",
                                                                                                                 ifelse(use_data()$TEAM=="St. Louis Rams","https://static.nfl.com/static/site/img/logos/svg/teams/LA.svg",
                                                                                                                        ifelse(use_data()$TEAM=="Jacksonville Jaguars","https://static.nfl.com/static/site/img/logos/svg/teams/JAX.svg",
                                                                                                                               ifelse(use_data()$TEAM=="Philadelphia Eagles","https://static.nfl.com/static/site/img/logos/svg/teams/PHI.svg",
                                                                                                                                      ifelse(use_data()$TEAM=="Seattle Seahawks","https://static.nfl.com/static/site/img/logos/svg/teams/SEA.svg",
                                                                                                                                             ifelse(use_data()$TEAM=="Oakland Raiders","https://static.nfl.com/static/site/img/logos/svg/teams/OAK.svg",
                                                                                                                                                    ifelse(use_data()$TEAM=="Dallas Cowboys","https://static.nfl.com/static/site/img/logos/svg/teams/DAL.svg",
                                                                                                                                                           ifelse(use_data()$TEAM=="Carolina Panthers","https://static.nfl.com/static/site/img/logos/svg/teams/CAR.svg",
                                                                                                                                                                  ifelse(use_data()$TEAM=="New England Patriots","https://static.nfl.com/static/site/img/logos/svg/teams/NE.svg",
                                                                                                                                                                         ifelse(use_data()$TEAM=="Indianapolis Colts","https://static.nfl.com/static/site/img/logos/svg/teams/IND.svg",
                                                                                                                                                                                ifelse(use_data()$TEAM=="Minnesota Vikings","https://static.nfl.com/static/site/img/logos/svg/teams/MIN.svg",
                                                                                                                                                                                       ifelse(use_data()$TEAM=="Washington Redskins","https://static.nfl.com/static/site/img/logos/svg/teams/WAS.svg",
                                                                                                                                                                                              ifelse(use_data()$TEAM=="Tennessee Titans","https://static.nfl.com/static/site/img/logos/svg/teams/TEN.svg",
                                                                                                                                                                                                     ifelse(use_data()$TEAM=="San Diego Chargers","https://static.nfl.com/static/site/img/logos/svg/teams/LAC.svg",
                                                                                                                                                                                                            ifelse(use_data()$TEAM=="Denver Broncos","https://static.nfl.com/static/site/img/logos/svg/teams/DEN.svg",
                                                                                                                                                                                                                   ifelse(use_data()$TEAM=="Detroit Lions","https://static.nfl.com/static/site/img/logos/svg/teams/DET.svg",
                                                                                                                                                                                                                          ifelse(use_data()$TEAM=="Green Bay Packers","https://static.nfl.com/static/site/img/logos/svg/teams/GB.svg",
                                                                                                                                                                                                                                 ifelse(use_data()$TEAM=="Arizona Cardinals","https://static.nfl.com/static/site/img/logos/svg/teams/ARI.svg",
                                                                                                                                                                                                                                        ifelse(use_data()$TEAM=="New York Giants", "https://static.nfl.com/static/site/img/logos/svg/teams/NYG.svg",
                                                                                                                                                                                                                                               ifelse(use_data()$TEAM=="New York Jets", "https://static.nfl.com/static/site/img/logos/svg/teams/NYJ.svg","")))))))))))))))))))))))))))))))), iconWidth = use_data()$NumSuspensions^1.5 +20, iconHeight = use_data()$NumSuspensions^1.5 + 60)
    
    map <- use_data() %>%
      leaflet() %>% 
      addTiles() %>% 
      setView(-95, 37.5,
              zoom = 4) %>% 
      addMarkers(icon = ~NFL_icons, label=(~NumSuspensions), lng=~(as.numeric(LONGITUDE)), lat=~(as.numeric(LATITUDE)), popup= ~TEAM)
    map
  })
  
}

# call to shinyApp
shinyApp(ui = ui, server = server)



```

      
\newpage
2. Publish your app.  Then, go to the GitHub discussion "Shiny Apps" and reply to the message with (1) the URL to your published Shiny app; and (2) a paragraph explaining what story your Shiny app is telling, and how the interactivity you created enhances the telling of that story.

> ANSWER: Do not include anything here.  The link to your app and the paragraph should be posted to the "Shiny Apps" discussion thread on GitHub.